Domain Specific Language for high performance computing

### 2024-01-25 paper/paper.tex
20240117
0122
20240125

% \documentclass[ams]{U-AizuGT}
\documentclass[twocolumn, a4j]{article}
% \usepackage{pifont}
% \usepackage{graphicx}
% \usepackage{cite}

% \twocolumn[
\bibliographystyle{ieice}
\author{Ryuki Hiwada}
% \studentid{s1280076}
% \supervisor{Naohito Nakasato}
% ]
\title{Domain Specific Language for high performance computing}
\begin{document}
\maketitle
\section{Abstract}
\section{Introduction}
\subsection{Background}
In astrophysics and astronomy, to numerically calculate the dynamical 
evolution of N particles interacting gravitationally, N-body simulations 
are required. Figure 1 shows the equation for interparticle interactions 
in N-body simulations. If the equation is naively computed, the time
complexity of calculation of interparticle interactions is 
\begin{math}O(N^2) \end{math}, where 
N is the number of particles. Therefore, parallelization is required to
speed up numerical simulations. To write a parallelized code for a 
numerical simulation, a user needs to understand the architecture of 
computer systems in  detail. If a parallelized code is automatically
generated by only describing the formulas and data of the numerical
simulation, the above problems are solved. To realize the parallelization,
we will develop Domain Specific Language \lparen DSL \rparen, which is a 
programming language specialized to a domain, for example, SQL and HTML.
\subsection{implemented DSL}

\subsection{parallelization}

\subsection{Aim of this paper}

\section{Method}
\subsection{Use Sympy for DSL development.}

>>>>

\documentclass[ams]{U-AizuGT}
\usepackage{pifont}
\usepackage{graphicx}
\usepackage{cite}
\usepackage{mathtools}
\usepackage{listings,jvlisting}

\lstset{
    frame=single,
    numbers=left,
    tabsize=2
}
\bibliographystyle{ieice}
\author{Ryuki Hiwada}
\studentid{s1280076}
\supervisor{Naohito Nakasato}

\title{Domain Specific Language for high performance computing}
\begin{document}
\maketitle
\section{Abstract}
\section{Introduction}
\subsection{Background}
In astrophysics and astronomy, to numerically calculate the dynamical 
evolution of N particles interacting gravitationally, N-body simulations 
are required. Figure 1 shows the equation for interparticle interactions 
in N-body simulations. If the equation is naively computed, the time
complexity of calculation of interparticle interactions is 
\begin{math}O(N^2) \end{math}, where 
N is the number of particles. Therefore, parallelization is required to
speed up numerical simulations. To write a parallelized code for a 
numerical simulation, a user needs to understand the architecture of 
computer systems in  detail. If a parallelized code is automatically
generated by only describing the formulas and data of the numerical
simulation, the above problems are solved. To realize the parallelization,
we will develop Domain Specific Language \lparen DSL \rparen, which is a 
programming language specialized to a domain, for example, SQL and HTML.
\subsection{parallelization}
SIMD is one of the categories in Flynn's taxonomy 1, related to computer arithmetic
 processing. Using SIMD, arithmetic operations can be applied simultaneously to
multiple data. In this paper, we compute and parallelize multiple particles. For example, 
as shown in Figure 1, the computation of acceleration for different particles is independent
of each other, allowing the process to be parallelized. As a result, for instance Assuming 
that the variables for particles are double-precision floating-point numbers, we can execute 
an operation on four elements simultaneously using the AVX2 instruction, it is possible to 
perform operations on four elements. Therefore, compared to non-parallelized code, there is
a potential to accelerate the computation by four times. We will describe the design of a 
DSL in the next section to generate such parallelized code from the formulation of particle-particle interactions.
\subsection{Overview of DSL}

We have created a Domain Specific Language \lparen DSL \rparen named \lparen DSL name \rparen. This language's 
processor can read code describing particle interaction formulas and generate code 
that accelerates these computations. The language allows for the definition of 
variables and the description of interaction formulas. In the variable definition 
section, it's possible to define variables such as the mass and position of 
particles, variables for storing results, and other necessary variables. For the formula
part, we have enabled calculations including basic arithmetic operations, sqrt, power 
functions, and computing the norm of vectors. From this description, a kernel function 
is generated, which can then be called and used. We have created a Domain Specific 
Language \lparen DSL \rparen named \lparen DSL name\rparen. This language's processor can read code describing particle 
interaction formulas and generate code that accelerates these computations. The language allows
for the definition of variables and the description of interaction formulas. In the variable
definition section, it's possible to define variables such as the mass and position of 
particles, variables for storing results, and other necessary variables. For the formula part, 
we have enabled calculations including basic arithmetic operations, sqrt, power functions, and 
computing the norm of vectors. From this description, a kernel function is generated, which can then be called and used.
\subsection{Aim of this paper}

\section{implementation}
We use Sympy, a library in Python for implementing  the implementation of our \lparen DSL\rparen .
SymPy supports various operations, and the formulas using these operations are internally treated as 
syntax trees. We leveraged this functionality to read formulas and generate code capable of parallel execution.
For parallelization, we used an instruction set named Advanced Vector Extensions 2 (AVX2). AVX2 is one of the 
extended instruction sets implemented in Intel's CPUs. The next section will explain how parallelization is carried out using AVX2.

\subsection{AVX2 for parallelization}
To execute SIMD operations, SIMD-specific registers are used, 
and multiple data are computed simultaneously to match the bit 
size of the registers. For example, in the case of CPUs that 
support AVX2 instructions, the registers are 256 bits, allowing 
four double-precision floating-point numbers to be stored and 
computed at the same time. To explicitly handle SIMD instructions
 in high-level programming languages like C and C++, intrinsic 
 functions are used. Additionally, to handle SIMD-specific registers, 
 specialized types are used. For C++, the file immintrin.h, 
 which is available as a standard in the language, is included to 
 handle these. As an example of parallelization, the paper explains
  the parallelization of the gravitational interaction formula shown 
  in Figure 1.

\subsection{design of (name of DSL)}

\begin{lstlisting}[language=c++]
#include <iostream>

int main(){
    std::cout << "Hello, world!" << std::endl;
    return 0;
}
\end{lstlisting}
\subsection{implementation of (name of DSL)}

\subsection{Use Sympy for DSL development}
